---
title: "Final_analysis"
author: "AnnaClaire Marley"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```

```{r}

# total_carbon: total discounted carbon (not per acre)
# discount_haul_chip: the discounted cost *per acre* of hauling chips
# relative_carb: discounted carbon per acre
# analysis on opt tie break and opt tie break nocc

```

```{r}
# CARB data

  # cum_disc_harvest -- cost per acre
  # cum_merch_merch == carbon per acre
  # cum_discount_carb = carbon per acre above ground living biomass
  # total_discount_carb = sum of the carb
  # total relative carb  = total_discount_carb - baseline 
  # total_relative_carb*acres
  # cum_disc_harvest = cost per acre 

```


```{r, include = FALSE}
# counties of all plots
  counties <- read_csv("plot_county.csv") %>% 
    select(ID, lat, lon, NAME)
#variant data and private or federal
  all_data <- read_csv("all_data.csv")
```

```{r, echo = FALSE}
# read in CARB and OG dataframes with no discounting
CARB_00 <- read_csv("carb_cpu_00.csv")
OG_00 <- read_csv("relative_carb_og_00.csv")
```

```{r, echo = FALSE}
# read in CARB and OG dataframes with discount rate of 5%
CARB_05 <- read_csv("carb_cpu_05.csv")
OG_05 <- read_csv("relative_carb_og_05.csv")
```

```{r, echo = FALSE}
#  to get the optimal 
price <- 200 ### this is the max we are willing to pay  per ton

optimal <- CARB_05 %>% 
    filter(total_carbon > 0 ) %>%     
    mutate(value = (price * total_carbon) - total_cost) %>% 
    group_by(ID) %>% 
    filter(value > 0 &
             value == max(value))

opt_tie_break <- optimal %>% 
  group_by(ID) %>% 
  sample_n(1) %>% 
  ungroup()
```

```{r, echo = FALSE, include=FALSE, error=TRUE}
#  to get the optimal without clear cutting
price <- 200 ### this is the max we are willing to pay  per ton

optimal_noCC <- CARB_05 %>% 
    filter(total_carbon > 0 ) %>%     
    filter(!rxpackage %in% c("032", "033")) %>% 
    mutate(value = (price * total_carbon) - total_cost) %>% 
    group_by(ID) %>% 
    filter(value > 0 &
             value == max(value))

opt_tie_break_nocc <- optimal_noCC %>% 
  group_by(ID) %>% 
  sample_n(1) %>% 
  ungroup()
```

```{r, echo = FALSE}
# combine counties with opti_tie_Break
opt_tie_break_counties <- left_join(counties, opt_tie_break, by = "ID") %>% 
  filter(biosum_cond_id != "NA")

```


# **Package Frequency**
## With clear cutting as an option 
```{r, echo = FALSE}
# group by package and count package occurance
package_count <- opt_tie_break %>% 
  group_by(rxpackage) %>% 
  tally()

package_count$rxpackage <- factor(package_count$rxpackage, levels = package_count$rxpackage[order(-package_count$n)])

# plot it
ggplot(package_count, aes(as.factor(rxpackage), n)) +
  geom_col() +
  labs(
    x = "Treatment Package ID", 
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```


```{r, echo = FALSE, include = FALSE}

## Without clear cutting as an option 

# group by package and count package occurance
package_count_nocc <- opt_tie_break_nocc %>% 
  group_by(rxpackage) %>% 
  tally()

package_count_nocc$rxpackage <- factor(package_count_nocc$rxpackage, levels = package_count_nocc$rxpackage[order(-package_count_nocc$n)])

# plot it
ggplot(package_count_nocc, aes(as.factor(rxpackage), n)) +
  geom_col() +
  labs(
    x = "Treatment Package ID", 
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```


# Compare carbon consequences of grow only vs thinning vs clear cut

```{r, echo=FALSE}
# assign names for different types of treatments
opt_tie_break_counties$treatment_type = 0
   opt_tie_break_counties$treatment_type = 
     ifelse(opt_tie_break_counties$rxpackage %in% c("033","032"), 
                                     "clear_cut", opt_tie_break_counties$treatment_type)
   
   opt_tie_break_counties$treatment_type = 
     ifelse(opt_tie_break_counties$rxpackage == "031", 
                                     "grow_only", opt_tie_break_counties$treatment_type)
   
   opt_tie_break_counties$treatment_type = 
     ifelse(!(opt_tie_break_counties$rxpackage %in% c("033","032", "031")), 
                                       "thinning", opt_tie_break_counties$treatment_type)

```

```{r, echo = FALSE}
### plot the differences ####

# boxplot of cost per unit carbon range
ggplot(opt_tie_break_counties, aes(x = treatment_type, y = cpu)) +
  geom_boxplot() +
  labs(x = "Treatment Type",
       y = "Cost per Ton of Carbon") +
  theme_bw()

# boxplot of cost per acre
cost_plot <- opt_tie_break_counties %>% 
  mutate(cpa = total_cost/acres) %>% 
  ggplot(aes(x = treatment_type, y = cpa)) +
    geom_boxplot() +
    labs(x = "Treatment Type",
         y = "Cost per Acre") +
    theme_bw()

# boxplot of carbon stored per acre
carbon_plot <- opt_tie_break_counties %>% 
  mutate(carbon_pa = total_carbon/acres) %>% 
  ggplot(aes(x = treatment_type, y = carbon_pa)) +
    geom_boxplot() +
    labs(x = "Treatment Type",
         y = "Carbon Stored per Acre") +
    theme_bw()
carbon_plot
```

## How much carbon stored in the forest vs products

```{r, echo = FALSE}
## cum_discount_decay = chips
## cum_discount_carb = discounted total stand carbon
## cum_discount_merch = cummulative discounted merchantable

portions <- opt_tie_break_counties %>% 
  select(ID, rxpackage, NAME, acres, 
         cum_discount_carb, cum_discount_merch,
         total_discount_carb, cum_discount_decay) %>% 
  mutate(ratio_stand_carb = cum_discount_carb/total_discount_carb) %>% 
  mutate(ratio_merch_carb = cum_discount_merch/total_discount_carb) %>% 
  mutate(ratio_chip_carb = cum_discount_decay/total_discount_carb)

```

```{r, echo = FALSE}
# create a boxplot showing the differences in spread in ratios between merchatnable and stand carbon in general
ratio_boxplot <- portions %>% 
  select(ratio_stand_carb, ratio_merch_carb, ratio_chip_carb) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Ratio to Total Carbon",
    x = ""
  ) +
 scale_x_discrete(labels = c('Chip Carbon', 'Merchantable Carbon','Stand Carbon')) +
  theme_bw()
ratio_boxplot

```

### without clear cutting as an option

```{r, echo = FALSE}
# opt_tie_break_nocc
```

```{r, echo = FALSE}
## cum_discount_decay = chips
## cum_discount_carb = discounted total stand carbon
## cum_discount_merch = cummulative discounted merchantable

portions_nocc <- opt_tie_break_nocc %>% 
  select(ID, rxpackage, acres, 
         cum_discount_carb, cum_discount_merch,
         total_discount_carb, cum_discount_decay) %>% 
  mutate(ratio_stand_carb = cum_discount_carb/total_discount_carb) %>% 
  mutate(ratio_merch_carb = cum_discount_merch/total_discount_carb) %>% 
  mutate(ratio_chip_carb = cum_discount_decay/total_discount_carb)

```

```{r, echo = FALSE}
# create a boxplot showing the differences in spread in ratios between merchatnable and stand carbon in general
ratio_boxplot <- portions_nocc %>% 
  select(ratio_stand_carb, ratio_merch_carb, ratio_chip_carb) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Ratio to Total Carbon",
    x = ""
  ) +
 scale_x_discrete(labels = c('Chip Carbon', 'Merchantable Carbon','Stand Carbon')) +
  theme_bw()
ratio_boxplot

```

## Allocation of costs per ton of carbon 
```{r, echo = FALSE}

# change the discounted costs to total discounted cost by multiplying by aces and then make cpu
costs <- opt_tie_break_counties %>% 
  select(ID, NAME, acres, rxpackage, cum_disc_harvest, 
         cum_disc_haul_chip, cum_disc_haul_merch, 
         total_discount_carb, total_carbon) %>% 
  mutate(harvest_cost = cum_disc_harvest*acres) %>% 
  mutate(haul_chip_cost = cum_disc_haul_chip*acres) %>% 
  mutate(haul_merch_cost = cum_disc_haul_merch*acres) %>% 
  mutate(cpu_harvest = harvest_cost/total_carbon) %>% 
  mutate(cpu_haul_chip = haul_chip_cost/total_carbon) %>% 
  mutate(cpu_haul_merch = haul_merch_cost/total_carbon)

costs_spread <- costs %>% 
  select(cpu_harvest, cpu_haul_chip, cpu_haul_merch) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Cost per Ton of Carbon",
    x = "Cost Type"
  ) +
  theme_bw()
  
costs_spread
```

## Allocation of costs per unit by counties
```{r, echo = FALSE}
costs_mean <- costs %>% 
  select(NAME, cpu_harvest, cpu_haul_chip, cpu_haul_merch) %>% 
  gather(cost_type, cost, cpu_harvest:cpu_haul_merch) %>% 
  group_by(NAME, cost_type) %>% 
  summarize(mean_cost = mean(cost),
            sd_cost = sd(cost)) %>% 
  filter(NAME != "NA")

# plot bar chart with error bars
ggplot(costs_mean, aes(x = reorder(NAME, mean_cost), y = mean_cost, fill = cost_type)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip() +
  labs(
    x = "County",
    y = "Average Cost per Ton of Carbon",
    fill = "Cost Type"
  ) +
  scale_fill_discrete(labels = c("Harvest", "Hauling chips", "Hauling merchantable")) +
  theme_bw()

```

## CPU and carbon stored across the different eco regions (variants)

```{r, echo = FALSE}

variants <- all_data %>% 
  select(ID, fvs_variant) %>% 
  group_by(ID, fvs_variant) %>% 
  summarize()

simple_plus <- opt_tie_break_counties %>% 
  left_join(variants, by = "ID")

```

### Total disocunted carbon per acre stored by variant

```{r, echo = FALSE}

variant_carbon <- simple_plus %>% 
  select(biosum_cond_id, ID, rxpackage, fvs_variant, owngrpcd, cpu,
         total_discount_carb, acres, total_cost) 

variant_boxplot <- variant_carbon %>% 
  ggplot(aes(x = fvs_variant, y = total_discount_carb)) +
  geom_boxplot() +
  labs(
    x = "FVS Variant",
    y = "Total Discounted Carbon per Acre"
  ) +
  theme_bw()
variant_boxplot

```

## CPU split by variant

```{r, echo = FALSE}

variant_cost_boxplot <- variant_carbon %>% 
  ggplot(aes(x = fvs_variant, y = cpu)) +
  geom_boxplot() +
  labs(
    x = "FVS Variant",
    y = "Cost ($) per Ton of Carbon"
  ) +
  scale_y_continuous(limits = c(-750, 200)) +
  theme_bw()

variant_cost_boxplot

```

# CPU for federal and private land

```{r, echo=FALSE}

fed_priv <- variant_carbon %>% 
  select(cpu, owngrpcd) %>% 
  mutate(ownership = ifelse(owngrpcd == 40, "Private", "Federal"))

ggplot(fed_priv, aes(x = ownership, y = cpu)) +
  geom_boxplot() +
  labs(
    x = "Ownership",
    y = "Cost per Ton of Carbon"
  ) + 
  scale_y_continuous(limits = c(-750, 200)) +
  theme_bw()
  

```


## visualization of differences in methods
```{r, include = FALSE}
all_data <- read_csv("all_data.csv")
source("discount_all_new.R")
source("old_discount_method.R") 

decay_pct <- 1
char_pct <- 0

#### decay rates for merch and non-merch
merch_decay <- read_csv("softwood_lumber_decay.csv")
non_merch <- read_delim("chip_pathways.txt", delim = ",")


  
test <- add_discounting_old(all_data, 34, "018")
base <- add_discounting_old(all_data, 34, "031")
  

### LEGEND:
# grow only is blue
# red/brown : merch and chip carbon
# green is total stand
# dashed: total overall carbon
ggplot(test, aes(x = time, y = cum_discount_carb)) +
  geom_line(color = "green") +
  geom_line(aes(x = time, y = cum_discount_merch), color = "red") +
  geom_line(aes(x = time, y = cum_discount_decay), color = "brown") +
  geom_line(aes(x = time, y = total_discount_carb), linetype = "dashed") +
  geom_line(data = base, aes(time, total_discount_carb), color = "blue") +
  labs(title = "old method")


## now repeat with same plot/ package using the new method:
df_18 <- add_discounting_new_long(all_data, 34, "018")
df_go <- add_discounting_new_long(all_data, 34, "031")



# same legend
ggplot(df_18, aes(x = time, y = cum_discount_carb)) +
  geom_line(color = "green") +
  geom_line(aes(x = time, y = cum_discount_merch), color = "red") +
  geom_line(aes(x = time, y = cum_discount_decay), color = "brown") +
  geom_line(aes(x = time, y = total_discount_carb), linetype = "dashed") +
  geom_line(data = df_go, aes(x = time, y = total_discount_carb), color = "blue") +
  labs(title = "new method")
```








