---
title: "Report Graphs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(sf)
library(ggspatial)
library(RColorBrewer)
library(car)
options(scipen=999)
```

```{r, include = FALSE}
# counties of all plots
  counties <- read_csv("plot_county.csv") %>% 
    select(ID, NAME)
#variant data and private or federal
  all_data <- read_csv("all_data.csv")
```

```{r, include = FALSE}
# read in CARB and OG dataframes with no discounting
CARB_00 <- read_csv("carb_cpu_00.csv")
OG_00 <- read_csv("relative_carb_og_00.csv")
```

```{r, include = FALSE}
# read in CARB and OG dataframes with discount rate of 5%
CARB_05 <- read_csv("carb_cpu_05.csv")
OG_05 <- read_csv("relative_carb_og_05.csv")
```

```{r, echo = FALSE}
#  function to get the optimal plots and break ties
# all_plots is the raw data of all of the lots
# price is the max we are willing to pay per ton
optimal_generator = function(all_plots, price = 200, growonly = "yes"){
  optimal <- all_plots %>% 
      filter(total_carbon > 0) 
    
  if (growonly == "no"){
    optimal <- optimal %>% 
      filter(rxpackage != "031") %>% 
      filter(cum_discount_cost != 0)
  } else{
    NA
  }
  
  optimal <- optimal %>% 
      mutate(value = (price * total_carbon) - total_cost) %>% 
      group_by(ID) %>% 
      filter(value > 0 &
               value == max(value))
  
  opt_tie_break <- optimal %>% 
    group_by(ID) %>% 
    sample_n(1) %>% 
    ungroup()
  
  return(opt_tie_break)
}
```

```{r, include = FALSE}
# genrate optimal and tie break for 5% discount rates
carb_05_opt <- optimal_generator(CARB_05) %>% 
  left_join(counties, by = "ID")
OG_05_opt <- optimal_generator(OG_05) %>% 
  left_join(counties, by = "ID")
```

```{r, include = FALSE}
# genrate optimal and tie break for 5% discount rates without grow only as an option
carb_05_opt_ng <- optimal_generator(CARB_05, growonly = "no") %>% 
  left_join(counties, by = "ID")
OG_05_opt_ng <- optimal_generator(OG_05, growonly = "no") %>% 
  left_join(counties, by = "ID")
```


```{r, include=FALSE}
# with no discounting
carb_00_opt <- optimal_generator(CARB_00) %>% 
  left_join(counties, by = "ID")
OG_00_opt <- optimal_generator(OG_00) %>% 
  left_join(counties, by = "ID")
```


## Compare most frequently chosen prescriptions
```{r, include = FALSE}
# function to graph most frequently chosen prescriptions
# optimal_df: the optimal dataframe from above
# base_type: the title to the graph to differentiate between CARB and OG
plot_common_rx = function(optimal_df, base_type = ""){
  # group by package and count package occurance
  package_count <- optimal_df %>% 
    group_by(rxpackage) %>% 
    tally()
  
  package_count$rxpackage <- factor(package_count$rxpackage, levels = package_count$rxpackage[order(-package_count$n)])
   
  # plot it
 plot <- ggplot(package_count, aes(as.factor(rxpackage), n)) +
    geom_col() +
    labs(
      x = "Treatment Package ID", 
      y = "Count",
      title = base_type
    ) +
    scale_y_continuous(expand = c(0,0)) +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5))
 
 return(plot)
 
}
```

#### with discount rate of 5%
```{r, echo = FALSE}
# with discounting
plot_common_rx(carb_05_opt, "CARB Baseline")
plot_common_rx(OG_05_opt, "Original Baseline")
```

#### with discount rate of 5% but without allowing for grow only
```{r, echo = FALSE}
# with discounting
plot_common_rx(carb_05_opt_ng, "CARB Baseline")
plot_common_rx(OG_05_opt_ng, "Original Baseline")
```


#### without discount rate 
```{r, echo = FALSE}
# with discounting
plot_common_rx(carb_00_opt, "CARB Baseline")
plot_common_rx(OG_00_opt, "Original Baseline")
```



## Compare carbon consequences of grow only vs thinning vs clear cut 
### with optimal run 

```{r, include = FALSE}
# get the dataframes ready and compine
carb_cpu <- carb_05_opt %>% 
  select(ID, acres, rxpackage, total_relative_carb, cum_discount_cost, cpu) %>% 
  rename(relative_carb = total_relative_carb) %>% 
  mutate(baseline_type = "CARB")
og_cpu <- OG_05_opt %>% 
  select(ID, acres, rxpackage, relative_carb, cum_discount_cost, cpu) %>% 
  mutate(baseline_type = "Original")
carb_og_cpu <- rbind(carb_cpu, og_cpu)
```

```{r, include = FALSE}
# assign names for different types of treatments
carb_og_cpu$treatment_type = 0
   carb_og_cpu$treatment_type = 
     ifelse(carb_og_cpu$rxpackage %in% c("033","032"), 
                                     "clear cut", carb_og_cpu$treatment_type)
   
   carb_og_cpu$treatment_type = 
     ifelse(carb_og_cpu$rxpackage == "031", 
                                     "grow only", carb_og_cpu$treatment_type)
   
   carb_og_cpu$treatment_type = 
     ifelse(!(carb_og_cpu$rxpackage %in% c("033","032", "031")), 
                                       "thinning", carb_og_cpu$treatment_type)
   
 carb_og_cpu <- carb_og_cpu %>% 
   filter(relative_carb <= 40)
```

```{r, echo = FALSE}
# boxplot of carbon stored per acre
carbon_plot <- carb_og_cpu %>% 
  ggplot(aes(x = treatment_type, y = relative_carb)) +
    geom_boxplot(aes(fill = baseline_type)) +
    labs(x = "Treatment Type",
         y = "Carbon Stored per Acre",
         fill = "Baseline Type") +
    theme_bw()
carbon_plot
```

#### with all plots
```{r, include = FALSE}
# get the dataframes ready and compine
carb_cpu_all <- CARB_05 %>% 
  select(ID, acres, rxpackage, total_relative_carb, cum_discount_cost, cpu) %>% 
  rename(relative_carb = total_relative_carb) %>% 
  mutate(baseline_type = "CARB")
og_cpu_all <- OG_05 %>% 
  select(ID, acres, rxpackage, relative_carb, cum_discount_cost, cpu) %>% 
  mutate(baseline_type = "BAU")
carb_og_cpu_all <- rbind(carb_cpu_all, og_cpu_all)
```

```{r, include = FALSE}
# assign names for different types of treatments
carb_og_cpu_all$treatment_type = 0
   carb_og_cpu_all$treatment_type = 
     ifelse(carb_og_cpu_all$rxpackage %in% c("033","032"), 
                                     "clear cut", carb_og_cpu_all$treatment_type)
   
   carb_og_cpu_all$treatment_type = 
     ifelse(carb_og_cpu_all$rxpackage == "031", 
                                     "grow only", carb_og_cpu_all$treatment_type)
   
   carb_og_cpu_all$treatment_type = 
     ifelse(!(carb_og_cpu_all$rxpackage %in% c("033","032", "031")), 
                                       "thinning", carb_og_cpu_all$treatment_type)
   
 carb_og_cpu_all <- carb_og_cpu_all %>% 
   filter(relative_carb <= 40)
```

```{r, echo = FALSE}
# boxplot of carbon stored per acre
carbon_plot_all <- carb_og_cpu_all %>% 
  ggplot(aes(x = treatment_type, y = relative_carb)) +
    geom_boxplot(aes(fill = baseline_type)) +
    labs(x = "Treatment Type",
         y = "Carbon Stored per Acre",
         fill = "Baseline Type") +
    theme_bw()
carbon_plot_all
```

### carbon stored per acre stats
```{r, echo = FALSE}
CARB_carb <- carb_og_cpu_all %>% 
  select(baseline_type, treatment_type, relative_carb) %>% 
  filter(baseline_type == "CARB")
# what percent of relative carbon is above 0?
# 2,286 grow only scenarios and 1542 were above 0 
# 54872 thinning scenarios and 26247 were above 0
  # 28625 we below zero
# 4574 clear cut scenarios and 2961 were below zero
carb_go <- CARB_carb %>% 
  filter(treatment_type == "clear cut") %>% 
  filter(relative_carb < 0)
# test if normally distributed
ggplot(CARB_carb, aes(x = relative_carb)) +
  geom_histogram(aes(fill = treatment_type), bins = 100) + 
  facet_wrap(~treatment_type) + #if want to make mismatching scale write scale = "free"
  theme_classic()
 ggplot(CARB_carb, aes(sample = relative_carb)) +
  geom_qq(aes(color = treatment_type)) +
  facet_wrap(~treatment_type) +
  scale_y_continuous(limits = c(-500,500))
```

```{r, echo = FALSE}
# test of variances
levene_test_CARB <- leveneTest(relative_carb ~ treatment_type, data= CARB_carb)
# there is a significant difference in variances between the data
anova_CARB <- aov(relative_carb ~ treatment_type, data= CARB_carb)
summary(anova_CARB)
# there is asignificant difference in carbon per acre between thinning treatments for the CARB data
```

```{r, include = FALSE}
BAU_carb <- carb_og_cpu_all %>% 
  select(baseline_type, treatment_type, relative_carb) %>% 
  filter(baseline_type == "BAU") 
# test if normally distributed
ggplot(BAU_carb, aes(x = relative_carb)) +
  geom_histogram(aes(fill = treatment_type), bins = 100) + 
  facet_wrap(~treatment_type) + #if want to make mismatching scale write scale = "free"
  theme_classic()
 ggplot(BAU_carb, aes(sample = relative_carb)) +
  geom_qq(aes(color = treatment_type)) +
  facet_wrap(~treatment_type) +
  scale_y_continuous(limits = c(-500,500))
```

```{r, echo = FALSE}
# test of variances
levene_test_BAU <- leveneTest(relative_carb ~ treatment_type, data = BAU_carb)
# there is asignificant difference in variances between the data
anova_BAU <- aov(relative_carb ~ treatment_type, data= BAU_carb)
summary(anova_BAU)
# there is a significant difference in carbon per acre between thinning treatments for the BAU data
TukeyHSD(anova_BAU)
```

```{r, echo = FALSE}
var.test(relative_carb ~ baseline_type, data = carb_og_cpu_all)
# variances are different 
t.test(relative_carb ~ baseline_type, data = carb_og_cpu_all, alternative = "less")

```

```{r, echo = FALSE}
# what percent of relative carbon is below 0?
# 4,578 clear cut scenarios and 2921 were below 0 
# 54936 thinning scenarios and 44987 were below 0
# 2288 grow only scenarios and 792 were above 0
bau_go <- BAU_carb %>% 
  filter(treatment_type == "grow only") 
bau_thin <- BAU_carb %>% 
  filter(treatment_type == "thinning") 
# ## within baseline t tests #####
# check equal variances
var.test(bau_go$relative_carb, bau_thin$relative_carb)
# they do not have equal variances 
# so need to do a wa
```


## CPU by baseline type
### CPU has been filtered to be above - $1000/ton carbon 
#### With optimal plots
```{r, echo = FALSE}
# boxplot of carbon stored per acre
cpu_plot <- carb_og_cpu %>% 
  filter(cpu >= -1000) %>% 
  filter(relative_carb > 0) %>% 
  ggplot(aes(x = treatment_type, y = cpu)) +
    geom_boxplot(aes(fill = baseline_type)) +
    labs(x = "Treatment Type",
         y = "Cost per Ton of Carbon",
         fill = "Baseline Type") +
    theme_bw()
cpu_plot
```

#### With all plots
```{r, echo = FALSE}
# boxplot of carbon stored per acre
cpu_plot_all <- carb_og_cpu_all %>% 
  filter(relative_carb >= 0) %>%
  filter(cpu >= -1000) %>% 
  filter(cpu < 1000) %>% 
  ggplot(aes(x = treatment_type, y = cpu)) +
    geom_boxplot(aes(fill = baseline_type)) +
    labs(x = "Treatment Type",
         y = "Cost per Ton of Carbon",
         fill = "Baseline Type") +
    theme_bw()
cpu_plot_all
```

## alocation of plots in public or private with no grow only allowed
```{r, echo = FALSE}
fed_priv_carb_ng <- carb_05_opt_ng %>% 
  select(ID, cpu, owngrpcd) %>% 
  mutate(ownership = ifelse(owngrpcd == 40, "Private", "Federal")) %>% 
  group_by(ownership) %>% 
  tally() %>% 
  mutate(type = "CARB")
fed_priv_og_ng <- OG_05_opt_ng %>% 
  select(ID, cpu, owngrpcd) %>% 
  mutate(ownership = ifelse(owngrpcd == 40, "Private", "Federal")) %>% 
  group_by(ownership) %>% 
  tally() %>% 
  mutate(type = "BAU")
### combine 
own_og_carb <- rbind(fed_priv_carb_ng, fed_priv_og_ng)
add_carb_0 <- data.frame(ownership = c("Private"), n = c(0), type = c("CARB"))
own_og_carb <- rbind(own_og_carb, add_carb_0)
```

```{r, echo = FALSE}
# graph it
ggplot(own_og_carb, aes(x = ownership, y = n)) +
  geom_col(aes(fill = type), position = "dodge") +
  scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Ownership Type",
    y = "Number of Plots Selected",
    fill = "Baseline"
  ) +
  theme_bw()
```






## Allocation of costs per unit by counties without grow only scenarios

### Original Baseline
```{r, include = FALSE}
# change the discounted costs to total discounted cost by multiplying by aces and then make cpu
costs <- OG_05_opt_ng %>% 
  select(ID, NAME, acres, rxpackage, cum_disc_harvest, 
         cum_disc_haul_chip, cum_disc_haul_merch, 
         total_discount_carb, relative_carb) %>% 
  mutate(cpu_harvest = cum_disc_harvest/relative_carb) %>% 
  mutate(cpu_haul_chip = cum_disc_haul_chip/relative_carb) %>% 
  mutate(cpu_haul_merch = cum_disc_haul_merch/relative_carb)
```

```{r, echo = FALSE}
costs_mean <- costs %>% 
  select(NAME, cpu_harvest, cpu_haul_chip, cpu_haul_merch) %>% 
  gather(cost_type, cost, cpu_harvest:cpu_haul_merch) %>% 
  group_by(NAME, cost_type) %>% 
  summarize(mean_cost = mean(cost),
            sd_cost = sd(cost)) %>% 
  filter(NAME != "NA")
```

```{r, echo = FALSE}
# plot bar chart with error bars
ggplot(costs_mean, aes(x = reorder(NAME, mean_cost), y = mean_cost, fill = cost_type)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip() +
  labs(
    x = "County",
    y = "Average Cost per Ton of Carbon",
    fill = "Cost Type"
  ) +
  scale_fill_manual(values = c("#138f23", "#ffa600", "#695746"), 
                    labels = c("Harvest", "Hauling chips", "Hauling merchantable")) +
  theme_bw()
```
### CARB Baseline
```{r, include = FALSE}
# change the discounted costs to total discounted cost by multiplying by aces and then make cpu
costs_carb <- carb_05_opt_ng %>% 
  select(ID, NAME, acres, rxpackage, cum_disc_harvest, 
         cum_disc_haul_chip, cum_disc_haul_merch, 
         total_discount_carb, total_relative_carb) %>% 
  mutate(cpu_harvest = cum_disc_harvest/total_relative_carb) %>% 
  mutate(cpu_haul_chip = cum_disc_haul_chip/total_relative_carb) %>% 
  mutate(cpu_haul_merch = cum_disc_haul_merch/total_relative_carb)
```

```{r, include = FALSE}
costs_mean_carb <- costs_carb %>% 
  select(NAME, cpu_harvest, cpu_haul_chip, cpu_haul_merch) %>% 
  gather(cost_type, cost, cpu_harvest:cpu_haul_merch) %>% 
  group_by(NAME, cost_type) %>% 
  summarize(mean_cost = mean(cost),
            sd_cost = sd(cost)) %>% 
  filter(NAME != "NA")
```

```{r, echo = FALSE}
# plot bar chart with error bars
ggplot(costs_mean_carb, aes(x = reorder(NAME, mean_cost), y = mean_cost, fill = cost_type)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 600)) +
  coord_flip() +
  labs(
    x = "County",
    y = "Average Cost per Ton of Carbon",
    fill = "Cost Type"
  ) +
  scale_fill_manual(values = c("#138f23", "#ffa600", "#695746"), 
                    labels = c("Harvest", "Hauling chips", "Hauling merchantable")) +
  theme_bw()
```

### carbon distribution

#### CARB Baseline
```{r, include = FALSE}
# make ratios of each type of carbon to total carbon 
CARB_spread_carb <- carb_05_opt_ng %>% 
  select(ID, NAME, acres, rxpackage, total_discount_carb,
         cum_discount_carb, cum_discount_merch, cum_discount_decay) %>% 
  mutate(cr_harvest = cum_discount_carb/total_discount_carb) %>% 
  mutate(cr_chip = cum_discount_decay/total_discount_carb) %>% 
  mutate(cr_merch = cum_discount_merch/total_discount_carb)
# create a new dataframe for graphing
carb_spread_clean <- CARB_spread_carb %>% 
  select(NAME, cr_harvest, cr_chip, cr_merch) %>% 
  gather(type, carbon, cr_harvest:cr_merch) %>% 
  group_by(NAME, type) %>% 
  summarize(mean_carbon = mean(carbon),
            sd_carbon = sd(carbon)) %>% 
  filter(NAME != "NA")
```

```{r, echo = FALSE}
ggplot(carb_spread_clean, aes(x = reorder(NAME, mean_carbon), y = mean_carbon, fill = type)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip() +
  labs(
    x = "County",
    y = "Average Carbon to Total Ratio",
    fill = "Type"
  ) +
  scale_fill_manual(values = c("#138f23", "#ffa600", "#695746"), 
                    labels = c("Chips", "Forest", "Merchantable")) +
  theme_bw()
```

#### BAU baselin
```{r, include = FALSE}
# make ratios of each type of carbon to total carbon 
BAU_spread_carb <- OG_05_opt_ng %>% 
  select(ID, NAME, acres, rxpackage, total_discount_carb,
         cum_discount_carb, cum_discount_merch, cum_discount_decay) %>% 
  mutate(cr_harvest = cum_discount_carb/total_discount_carb) %>% 
  mutate(cr_chip = cum_discount_decay/total_discount_carb) %>% 
  mutate(cr_merch = cum_discount_merch/total_discount_carb)
# create a new dataframe for graphing
bau_spread_clean <- BAU_spread_carb %>% 
  select(NAME, cr_harvest, cr_chip, cr_merch) %>% 
  gather(type, carbon, cr_harvest:cr_merch) %>% 
  group_by(NAME, type) %>% 
  summarize(mean_carbon = mean(carbon),
            sd_carbon = sd(carbon)) %>% 
  filter(NAME != "NA")
```

```{r, echo = FALSE}
ggplot(bau_spread_clean, aes(x = reorder(NAME, mean_carbon), y = mean_carbon, fill = type)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip() +
  labs(
    x = "County",
    y = "Average Carbon to Total Ratio",
    fill = "Type"
  ) +
  scale_fill_manual(values = c("#138f23", "#ffa600", "#695746"), 
                    labels = c("Chips", "Harvest", "Merchantable")) +
  theme_bw()
```

# MAPS

```{r, include = FALSE}
plot_loc_geom <- read_sf("spatial_data", layer = "plot_loc")
ca_counties <- read_sf("spatial_data/CA_Counties", layer = "CA_Counties_TIGER2016")
```

```{r, include = FALSE}
ca_counties_clean <- ca_counties %>% 
  select(NAME)
carb_05_cpu_ng <- carb_05_opt_ng %>% 
  select(ID, cpu)
OG_05_cpu_ng <- OG_05_opt_ng %>% 
  select(ID, cpu)
plots_cpu_carb <- plot_loc_geom %>% 
  select(ID) %>% 
  left_join(carb_05_cpu_ng) %>% 
  filter(cpu != "NA")
plots_cpu_og <- plot_loc_geom %>% 
  select(ID) %>% 
  left_join(OG_05_cpu_ng) %>% 
  filter(cpu != "NA")
# range of pcu is 0 - 200
# 0 - 66
```

```{r, echo=FALSE}
pal <- brewer.pal(5, "OrRd") # we select 7 colors from the palette
class(pal)
# plot(ca_counties_clean)
#plot(plots_cpu_carb["ID"])
```
## CARB baseline 
```{r, echo = FALSE}
ggplot(ca_counties_clean) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = plots_cpu_carb, aes(color = cpu), alpha = 0.5) +
  scale_color_gradient(low = "#ffeda0", high = "#f03b20") +
  theme_bw()
```

### BAU baselin
```{r, echo = FALSE}
plots_cpu_og_filt <- plots_cpu_og %>% 
  filter(cpu >= -200)
ggplot(ca_counties_clean) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = plots_cpu_og_filt, aes(color = cpu), alpha = 0.5) +
  scale_color_gradient(low = "#e5f5f9", high = "#0d3804") +
  theme_bw()
```