---
title: "Results_analysis"
author: "AnnaClaire Marley"
date: "2/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```

```{r, include = FALSE}

opt_tie_break <- read_csv("opt_tie_break.csv")
selected_sites <- read_csv("selected_sites.csv")

```


# **Package Frequency**
```{r, echo = FALSE}
# group by package and count package occurance
package_count <- opt_tie_break %>% 
  group_by(rxpackage) %>% 
  tally()

package_count$rxpackage <- factor(package_count$rxpackage, levels = package_count$rxpackage[order(-package_count$n)])

# plot it
ggplot(package_count, aes(as.factor(rxpackage), n)) +
  geom_col() +
  labs(
    title = "counts of optimal packages",
    x = "Treatment Package ID", 
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```


# **County level analsyes**

Average cost per acre by county
```{r, echo = FALSE}
## decay = chips
## cum_discount_carb = discounted total stand carbon
## cum_discount_merch = cummulative discounted merchantable
simple_acre <- selected_sites %>%
  select(ID, county) %>% 
  left_join(opt_tie_break, by = "ID") %>% 
  select(biosum_cond_id, ID, county, acres, rxpackage, 
         cum_discount_carb, cum_discount_merch, cum_discount_decay,
         total_cost, total_carbon, total_discount_carb, cpu) %>% 
  filter(biosum_cond_id != "NA")

cost_acre <- simple_acre %>% 
  mutate(cpa = total_cost/acres) %>% 
  group_by(county) %>% 
  summarize(mean_cpa = mean(cpa),
            sd_cpa = sd(cpa)) %>% 
  arrange(-mean_cpa)

```

Average carbon stored per acre by county
```{r, echo = FALSE}
carbon_acre <- simple_acre %>% 
 mutate(carb_pa = total_carbon/acres) %>% 
  group_by(county) %>% 
  summarize(mean_carb_pa = mean(carb_pa),
            sd_carb_pa = sd(carb_pa)) %>% 
  arrange(-mean_carb_pa)
```



# Compare carbon consequences of surface fuels method

NOTE: 32 and 33 are clear cut
31 = grow only
```{r, echo = FALSE}
# group the thinning treatments by surface fuels method
rx_fire = c("001", "005", "010", "012", "015", "020", "024")
pile_burn = c("002", "013")
lop_scatter = c("004", "006", "007", "008", "009", "011", "014", 
                "017", "018", "019", "023", "025", "026", "027")
masticate = c("003")

# assign them to the rxpackages
simple_acre$surface_fuel = 0
   simple_acre$surface_fuel = ifelse(simple_acre$rxpackage %in% rx_fire, 
                                     "rx_fire", simple_acre$surface_fuel)
   
   simple_acre$surface_fuel = ifelse(simple_acre$rxpackage %in% pile_burn, 
                                     "pile_burn", simple_acre$surface_fuel)
   
   simple_acre$surface_fuel = ifelse(simple_acre$rxpackage %in% lop_scatter, 
                                     "lop_scatter", simple_acre$surface_fuel)
   
   simple_acre$surface_fuel = ifelse(simple_acre$rxpackage %in% masticate,
                                     "masticate", simple_acre$surface_fuel)

  
# order it

surface_fuel_df <- simple_acre %>% 
  filter(surface_fuel != 0) %>% 
  mutate(carb_pa = total_carbon/acres)

# graph it
ggplot(surface_fuel_df, aes(x = surface_fuel, y = carb_pa)) +
  geom_boxplot() +
  labs(
    x = "Surface Fuels Method",
    y = "Total Carbon Stored per acre"
  )

```

# Compare carbon consequences of grow only vs thinning vs clear cut

```{r, echo=FALSE}
# assign names for different types of treatments
simple_acre$treatment_type = 0
   simple_acre$treatment_type = ifelse(simple_acre$rxpackage %in% c("033","032"), 
                                     "clear_cut", simple_acre$treatment_type)
   
   simple_acre$treatment_type = ifelse(simple_acre$rxpackage == "031", 
                                     "grow_only", simple_acre$treatment_type)
   
   simple_acre$treatment_type = ifelse(!(simple_acre$rxpackage %in% c("033","032", "031")), 
                                       "thinning", simple_acre$treatment_type)

```

```{r, echo = FALSE}
### plot the differences ####

# boxplot of cost per unit carbon range
ggplot(simple_acre, aes(x = treatment_type, y = cpu)) +
  geom_boxplot() +
  labs(x = "Treatment Type",
       y = "Cost per Ton of Carbon") +
  theme_bw()

# boxplot of cost per acre
cost_plot <- simple_acre %>% 
  mutate(cpa = total_cost/acres) %>% 
  ggplot(aes(x = treatment_type, y = cpa)) +
    geom_boxplot() +
    labs(x = "Treatment Type",
         y = "Cost per Acre") +
    theme_bw()

cost_plot

# boxplot of carbon stored per acre
carbon_plot <- simple_acre %>% 
  mutate(carbon_pa = total_carbon/acres) %>% 
  ggplot(aes(x = treatment_type, y = carbon_pa)) +
    geom_boxplot() +
    labs(x = "Treatment Type",
         y = "Carbon Stored per Acre") +
    theme_bw()
carbon_plot
```

Look at what the spread is between how much carbon is stored in products vs in the stand

```{r, echo = FALSE}
## decay = chips
## cum_discount_carb = discounted total stand carbon
## cum_discount_merch = cummulative discounted merchantable

portions <- simple_acre %>% 
  select(ID, rxpackage, county, acres, 
         cum_discount_carb, cum_discount_merch,
         total_discount_carb) %>% 
  mutate(ratio_stand_carb = cum_discount_carb/total_discount_carb) %>% 
  mutate(ratio_merch_carb = cum_discount_merch/total_discount_carb)

```

create a boxplot showing the differences in spread in ratios between merchatnable and stand carbon in general

```{r, echo = FALSE}

ratio_boxplot <- portions %>% 
  select(ratio_stand_carb, ratio_merch_carb) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Ratio to Total Carbon",
    x = ""
  ) +
  scale_x_discrete(labels = c('Merchantable Carbon','Stand Carbon'))
ratio_boxplot

```

same plot but just for clear cutting scenarios
NOTE: 32 and 33 are clear cut
31 = grow only
```{r, echo = FALSE}

ratio_boxplot_clearcut <- portions %>% 
  filter(rxpackage == "032" | rxpackage == "033") %>% 
  select(ratio_stand_carb, ratio_merch_carb) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Ratio to Total Carbon",
    x = ""
  ) +
  scale_x_discrete(labels = c('Merchantable Carbon','Stand Carbon'))
ratio_boxplot_clearcut

```

same plot but just for thinning scenarios

```{r, echo = FALSE}

ratio_boxplot_thinning <- portions %>% 
  filter(rxpackage != "032" | rxpackage != "033") %>% 
  select(ratio_stand_carb, ratio_merch_carb) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Ratio to Total Carbon",
    x = ""
  ) +
  scale_x_discrete(labels = c('Merchantable Carbon','Stand Carbon'))
ratio_boxplot_thinning

```

Look at how carbon stored varies across the different eco regions (variants)

```{r, echo = FALSE}
all_data <- read_csv("all_data.csv")

variants <- all_data %>% 
  select(ID, fvs_variant, owngrpcd) %>% 
  group_by(ID, fvs_variant, owngrpcd) %>% 
  summarize()

simple_plus <- simple_acre %>% 
  left_join(variants, by = "ID")

```

Total carbon stored by variant

```{r, echo = FALSE}

variant_carbon <- simple_plus %>% 
  select(biosum_cond_id, ID, rxpackage, fvs_variant, owngrpcd, cpu, 
         total_discount_carb, acres, total_cost) 

variant_boxplot <- variant_carbon %>% 
  ggplot(aes(x = fvs_variant, y = total_discount_carb)) +
  geom_boxplot() +
  labs(
    x = "FVS Variant",
    y = "Total Discounted Carbon"
  )
variant_boxplot

```

total cpu split by variant

```{r, echo = FALSE}

variant_cost_boxplot <- variant_carbon %>% 
  ggplot(aes(x = fvs_variant, y = cpu)) +
  geom_boxplot() +
  labs(
    x = "FVS Variant",
    y = "Cost ($) per Ton of Carbon"
  )

variant_cost_boxplot

```

cost per acre split by variant

```{r, echo = FALSE}
variant_cpa <- variant_carbon %>% 
  mutate(cpa = total_cost/acres) %>% 
  ggplot(aes(x = fvs_variant, y = cpa)) +
  geom_boxplot() +
  labs(
    x = "FVS Variant",
    y = "Cost per Acre"
  )
variant_cpa
```

Now let's look at the relationship between cost and elevation 

```{r, echo = FALSE}

plot_loc <- read_csv("plot_loc.csv")

elevation_variants <- plot_loc %>% 
  select(ID, elev) %>% 
  left_join(variant_carbon, by = "ID") %>% 
  filter(biosum_cond_id != "NA") %>% 
  mutate(cpa = total_cost/acres)

ggplot(elevation_variants, aes(x = elev, y = cpa)) +
  geom_point(aes(color = fvs_variant))
  

```

Seems like there's actually not a relationship between elevationa nd cost 


```{r, echo= FALSE}


```


## Allocation of costs
```{r, echo = FALSE}
costs <- opt_tie_break %>% 
  select(ID, acres, rxpackage, discount_harvest, 
         discount_haul_chip, discount_haul_merch, total_discount_carb) %>% 
  mutate(cpa_harvest = discount_harvest/acres) %>% 
  mutate(cpa_haul_chip = discount_haul_chip/acres) %>% 
  mutate(cpa_haul_merch = discount_haul_merch/acres)

costs_spread <- costs %>% 
  select(cpa_harvest, cpa_haul_chip, cpa_haul_merch) %>% 
  gather() %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  labs(
    y = "Cost per Acre",
    x = "Cost Type"
  )
  
costs_spread
```

